---
title: "Bike Sales part 1"
author: "Eddy"
date: "2024-10-11"
output: word_document
---

setting working directory
```{r}
getwd()
```
```{r}
#loading required packages
library(tidyverse)
library(magrittr)
```

loading the sales dataset
```{r}
#importing data
sales <- read.csv("D:/Programming/Datasets/Sales.csv")
sales
```

viewing the structure of the sales data set
```{r}
#structure
str(sales)
```
summary of sales data
```{r}
summary(sales)
```
checking for missing values in the dataset
```{r}
sum(is.na(sales))
```
checking for duplicates
```{r}
duplicates <- sales[duplicated(sales),]
duplicates
```
removing duplicates
```{r}
sales <- sales[!duplicated(sales),]
```

Analysis
What's the mean of `Customers_Age`?
```{r}
#mean of customer age
mean(sales$Customer_Age)
```
Visualization of density plot and box plot of customer age
```{r}
#density plot of customer age
ggplot(sales, aes(x = Customer_Age)) +
  geom_density(fill = "skyblue", alpha = 0.5) +
  labs(title = "Density Plot of Customer Age",
       x = "Customer Age",
       y = "Density")
```

```{r}
#boxplot of customer age
ggplot(sales, aes(y = Customer_Age)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Boxplot of Customer Age",
       y = "Customer Age")
```
Mean of Order quantity
```{r}
mean(sales$Order_Quantity)
```
Visualization of order quantity
```{r}
ggplot(sales, aes(x = Order_Quantity)) +
  geom_histogram(fill = "lightblue") +
  labs(title = "Histogram of Order Quality",
       x = "Order Quality",
       y = "Frequency")
```
```{r}
#boxplot of order quality
ggplot(sales, aes(y=Order_Quantity)) +
  geom_boxplot(fill="lightgreen") +
  labs(title = "Boxplot of order Quality",
       y = "Order Quality")
```
Number of sales per year
```{r}
#yearly sales
yearly_sales <- sales %>% 
  group_by(Year) %>% 
  summarize(total_sales = sum(Order_Quantity)) %>% 
  arrange(Year)
yearly_sales
```
Pie chart of yearly sales
```{r}
ggplot(yearly_sales, aes(x = "", y = total_sales, fill = Year)) +
  geom_bar(width = 1, stat = "identity") + 
  coord_polar(theta = "y") +
  labs(title = "Pie Chart of Yearly Sales") +
  theme_void()
```

Sales per Month
```{r}
#monthly sales
monthly_sales <- sales %>% 
  group_by(Month) %>% 
  summarize(total_monthly_sales = sum(Order_Quantity)) %>% 
  arrange(Month)
monthly_sales
  
```

Bar plot of Monthly Sales
```{r}
ggplot(monthly_sales, aes(x = Month, y = total_monthly_sales)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Bar Plot of Monthly Sales", x = "Month", y = "Total Monthly Sales (Revenue)") +
  theme_minimal()
```
Which country has most sales
```{r}
most_sales_country <- sales %>% 
  select(Country, Order_Quantity) %>% 
  arrange(desc(Order_Quantity))
most_sales_country
```
Bar plot of sales per country
```{r}
ggplot(most_sales_country, aes(x= Country, y= Order_Quantity)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  labs(title = "Bar plot of Most saler per Country",
       x = "Country",
       y = "Country Sales")
```

Creating a list of unique products
```{r}
unique_products <- sales %>% 
  distinct(Product)
unique_products
```
Most sold product
```{r}
top_10_sold_product <- sales %>% 
  group_by(Product) %>% 
  summarize(product_sales = sum(Order_Quantity)) %>% 
  arrange(desc(product_sales)) %>% 
  head(10)
top_10_sold_product
```
Bar plot of top 10 sold products
```{r}
ggplot(top_10_sold_product, aes(x= Product, y= product_sales )) +
  geom_bar(stat = "Identity", fill= "lightgreen") +
  labs(title = "Bar Plot of Top Ten Products",
       y= "Product Sales",
       x = "Product") +
   theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#Relationship between Unit cost and Unit Price
```{r}
#correlation between unit cost and unit price
correlation <- cor(sales$Unit_Cost, sales$Unit_Price)
print(paste("The correlation is:", correlation))
```
There is high correlation of 0.99 between unit cost and unit price
An increase in unit cost increase the unit price also

Plotting a scatter plot to visualize the correlation
```{r}
ggplot(sales, aes(x= Unit_Cost, y= Unit_Price)) +
  geom_point() +
  geom_smooth(method = "lm", se=FALSE) +
  labs(title = "Scatter plot of Unit Cost against Unit Sales",
       x= "Unit Cost",
       y= "Unit Sales")
```
Relationship between Profit and Country
```{r}
#boxpot of profit by country
ggplot(sales, aes(x= Country, y=Profit, fill = Country)) +
  geom_boxplot() +
  labs(title = "Boxplot of Profit by Country",
       y = "Profit",
       x= "Country") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Relationship between customer age and country
```{r}
ggplot(sales, aes(x= Country, y= Customer_Age, fill= Country)) +
  geom_boxplot() +
  labs(title = "Boxplot of Customer Age by Country",
       x = "Country",
       y = "Customer Age") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```
Cleaning the date column
converting it from character to date type
```{r}
sales <- sales %>% 
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
str(sales$Date)
```
Number of orders made in France or Canada
```{r}
canada_france_orders <- sales %>% 
  group_by(Country) %>% 
  filter(Country %in% c("France", "Canada")) %>% 
  summarise(total_orders = n()) 
canada_france_orders
```
How many Bike Racks orders were made in Canada
```{r}
bike_rack_orders <- sales %>% 
  group_by(Country) %>% 
  filter(Country == "Canada" & Sub_Category == "Bike Racks") %>% 
  summarise(total_racks_orders = n())
bike_rack_orders
```

How many orders were made in each state of France
```{r}
france_states_order <- sales %>% 
  group_by(State) %>% 
  filter(Country == "France") %>% 
  summarise(france_states_total_sales = n()) %>% 
  arrange(desc(france_states_total_sales))
france_states_order
```
Bar plot visualization of sales in each state in France

```{r}
ggplot(france_states_order, aes(x=State, y= france_states_total_sales)) +
  geom_bar(stat = "identity", fill= "lightgreen") +
  labs(title = "Bar Plot of Sales in Each State in France",
       x= "States",
       y= "Total Sales") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Sales made per category
```{r}
product_sales_per_category <- sales %>% 
  group_by(Product_Category) %>% 
  summarise(category_sales = n())
product_sales_per_category
```

Pie chart of the result
```{r}
ggplot(product_sales_per_category, aes(x= "", y = category_sales, fill = Product_Category)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  labs(title = "Pie Chart of Product Sales Per Category") +
  theme_void()
```
Sales made per accessory sub category
```{r}
accessory_sales_sub_category <- sales %>% 
  group_by(Sub_Category) %>% 
  summarize(sub_category_sales = n())
accessory_sales_sub_category
```

Bar plot of the results
```{r}
ggplot(accessory_sales_sub_category, aes(x= Sub_Category, y= sub_category_sales)) +
  geom_bar(stat= "identity", fill = "brown") +
  labs(title = "Bar plot of Accessories Sales Per Sub Category",
       x= "Sub Category",
       y= "SubCategory Sales") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Number of orders per bike sub category
```{r}
bike_orders <- sales %>% 
  group_by(Sub_Category) %>% 
  filter(Product_Category == "Bikes") %>% 
  summarise(total_orders = n()) %>% 
  arrange(desc(total_orders))
bike_orders
```
Pie Chart of bike orders per category
```{r}
ggplot(bike_orders, aes(x= "", y= total_orders, fill=Sub_Category)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  labs(title = "Pie chart of Orders per Bike") +
  theme_void()
```
Which gender has the most amount of sales
```{r}
gender_sales <- sales %>% 
  group_by(Customer_Gender) %>% 
  summarise(total_sales = n()) %>% 
  arrange(desc(total_sales))
gender_sales
```

Sales with more than 500 in revenue made by men
```{r}
men_sales_over_500 <- sales %>% 
  #group_by(Customer_Gender, Revenue) %>% 
  filter(Customer_Gender == "M" & Revenue > 500) %>% 
  summarise(total_sales = sum(Revenue))

men_sales_over_500
```

Top 5 sales with the highest revenue
```{r}
Top_5_sales <- sales %>% 
  arrange(desc(Revenue)) %>% 
  head(5)
Top_5_sales
```

Sale with the highest revenue
```{r}
highest_sale <- sales %>% 
  arrange(desc(Revenue)) %>% 
  head(1)
highest_sale
```
Mean order quantity with order more than 10k in revenue
```{r}
more_10k_order <- sales %>% 
  filter(Revenue > 10000) %>% 
  summarise(mean_order = mean(Order_Quantity))
mean_10k_order
```
Mean order quality with order less than 10k in revenue
```{r}
less_10k_order <- sales %>% 
  filter(Revenue < 10000) %>% 
  summarize(mean_order = mean(Order_Quantity))
less_10k_order
```
Orders made in May 2016
```{r}
may_2016_orders <- sales %>% 
  filter(Month == "May" & Year == 2016) %>% 
  summarize(total_orders =n())
may_2016_orders
```
Orders between May and July 2016
```{r}
may_july_orders <- sales %>%
  filter(Year == 2016 & Month %in% c("May", "June", "July")) %>% 
  summarize(total_orders = n())
may_july_orders
```

Grouped boxplot per month with profit values
```{r}
# Create grouped box plot of Profit per Month
ggplot(sales, aes(x = factor(Month), y = Profit, fill = factor(Month))) +
  geom_boxplot() +
  labs(title = "Box Plot of Profit by Month", x = "Month", y = "Profit") +
  scale_x_discrete(labels = month.name[1:12]) +  # Display full month names
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
Add 7.2% tax on unit price of United States
```{r}
tax_us <- sales %>% 
  group_by(Unit_Price) %>% 
  filter(Country == "United States") %>% 
  summarize(increase_unit_price = 7.2 * Unit_Price) %>% 
  arrange(desc(increase_unit_price))
tax_us
```

