---
title: "Trade Analysis"
author: "Eddy"
date: "2024-10-11"
output: word_document
---
#Working on the Amazon dataset
#working directory
```{r}
getwd()
```
#loading required packages
```{r}
library(tidyverse)
library(magrittr)
```

#loading dataset
```{r}
amazon <- read.csv("D:/Programming/Datasets/amazon.csv")
head(amazon)
```

#structure of the data
```{r}
str(amazon)
```
#summary of the dataset
```{r}
summary(amazon)
```
#checking for missing values 
```{r}
sum(is.na(amazon))
```
#checking for duplicates
```{r}
duplicates <- amazon[duplicated(amazon),]
duplicates
```


#converting numerical columns from character to numeric
#convering the discounted price variable
```{r}
amazon <- amazon %>% 
  mutate(discounted_price_cleaned = as.numeric(gsub("[^0-9.]", "", discounted_price)))
summary(amazon$discounted_price_cleaned)
```
#converting the actual price variable
```{r}
amazon <- amazon %>% 
  mutate(actual_price_cleaned = as.numeric(gsub("[^0-9.]", "", actual_price)))
summary(amazon$actual_price_cleaned)
```
#converting discount percentage column
```{r}
amazon <- amazon %>% 
  mutate(discount_percentage_cleaned = as.numeric(gsub("%", "",
                                                       discount_percentage)))
```

#converting the rating column
```{r}
amazon <- amazon %>% 
  mutate(rating_cleaned = as.numeric(gsub("[^0-9.]", "", rating)), 
      rating_is_na = is.na(rating_cleaned))
summary(amazon$rating_cleaned)
```
#dealing with the NA created
```{r}
na_rows <- amazon %>% 
  filter(rating_is_na)
na_rows
```

#cleaning the rating_cleaned again
```{r}
amazon <- amazon %>% 
  mutate(rating_cleaned1 = gsub("\\|", "", rating_cleaned),
         rating_cleaned = as.numeric(gsub("[^0-9]", "", rating_cleaned1)))
summary(amazon$rating_cleaned)
```
#dropping the row with the missing
```{r}
amazon <- amazon %>% 
  filter(!is.na(rating_cleaned))
summary(amazon$rating_cleaned)
```
#converting the rating discount var
```{r}
amazon <- amazon %>% 
  mutate(rating_count_cleaned = as.numeric(gsub("[^0-9].", "", rating_count)))
summary(amazon$rating_count_cleaned)
```
#removing rows with NAs
```{r}
amazon <- amazon %>% 
  filter(!is.na(rating_count_cleaned))
summary(amazon$rating_count_cleaned)
```
#checking the variables
```{r}
names(amazon)
```
#Analysis
#1.Which products have the highest customer satisfaction based on ratings?
```{r}
product_satisfaction <- amazon %>% 
  group_by(product_id, product_name) %>% 
  summarize(
    average_rating = mean(rating_cleaned),
    total_ratings = sum(rating_count_cleaned)
  ) %>% 
  arrange(desc(average_rating))
```
#Top 10 products with the highest customer satisfaction
```{r}
top_10_products <- product_satisfaction %>% 
  top_n(10, average_rating)
top_10_products
```

#2.Analyze the products with the highest discount percentages. 
```{r}
top_discounted_products <- amazon %>%
  select(product_id, product_name, discount_percentage_cleaned) %>%
  arrange(desc(discount_percentage_cleaned)) %>%
  slice_head(n = 10) 
top_discounted_products
```
#3.Which product categories are most popular, and how do they perform in terms of customer ratings?
```{r}
category_performance <- amazon %>% 
  group_by(category) %>% 
  summarize(ratings_total = sum(rating_count_cleaned),
            ratings_mean = mean(rating_cleaned)) %>% 
  arrange(desc(ratings_total))
category_performance
```


#4.Is there a correlation between discounted price and product rating? 
```{r}
correlation <- cor(amazon$discount_percentage_cleaned, amazon$rating_cleaned)
print(paste("Correlation coefficiency:", correlation))
```
#very weak correlation
#discounted price does not significantly impact customer ratings.

#visualizing the correlation
```{r}
ggplot(amazon, aes(x= discounted_price_cleaned, y= rating_cleaned)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(title = "Relationship Between Discounted Price and Product Rating",
       x = "Discounted Price",
       y = "Product Rating")
```